---
title: "Deviance-Formulas"
output: github_document
---

This notebook shows the deviance formula for several different distributions.

# Load data

Load the dataCar dataset from the insuranceData package. This will provide a dataset we can use to build both a poisson and gamma glm.

```{r setup, include = TRUE}
library(insuranceData)
data(dataCar)
```

```{r}
head(dataCar)
```

# Poisson deviance

Add the log of the exposure (period of cover) to the data, so it can be used as an offset in a poisson glm to model claims frequency (i.e. # claims per exposure).

```{r}
dataCar$log_exposure <- log(dataCar$exposure)
```

## Claims frequency model

This models number of claims per exposure, where exposure is different for each records.

```{r}
frequency_model <- glm(formula = numclaims ~ veh_body + veh_age + gender + area + agecat + offset(log_exposure),
                       family = poisson(link = log),
                       data = dataCar)
```

```{r}
frequency_model
```

## Claims models

This model predict number of claims, note each record has the same exposure.

```{r}
claims_model <- glm(formula = numclaims ~ veh_body + veh_age + gender + area + agecat,
                    family = poisson(link = log),
                    data = dataCar)
```

```{r}
claims_model
```

## Poisson deviance formula

Deviance formulas can be found here;  http://docs.h2o.ai/h2o/latest-stable/h2o-docs/data-science/glm.html. Note that h2o multiplies the deviance by -1.

```{r}
poisson_deviance <- function(y, p) {
  
  d <- 2 * ifelse(y == 0, -(y - p), (y * log(y / p)) - (y - p))
  
  dev <- sum(d)
  
  return(dev)
  
}
```

## Compare to results from glm()

### Claims frequecny model

```{r}
poisson_deviance(dataCar$numclaims, frequency_model$fitted.values)
```

```{r}
frequency_model$deviance
```

```{r}
poisson_deviance(dataCar$numclaims, 
                 dataCar$exposure * sum(dataCar$numclaims) / sum(dataCar$exposure))
```


```{r}
frequency_model$null.deviance
```

### Claims count model

```{r}
poisson_deviance(dataCar$numclaims, claims_model$fitted.values)
```

```{r}
claims_model$deviance
```

```{r}
poisson_deviance(dataCar$numclaims, 
                 sum(dataCar$numclaims) / length(dataCar$numclaims))
```

```{r}
claims_model$null.deviance
```

# Gamma deviance

Add the average claims cost that can be used as the response in a claims cost model.

```{r}
dataCar$average_claimcst0 <- dataCar$claimcst0 / dataCar$numclaims
```

## Claims severity model

This models cost (severity) of a claim.

```{r}
severity_model <- glm(formula = average_claimcst0 ~ veh_body + veh_age + gender + area + agecat,
                      family = Gamma(link = log),
                      data = dataCar,
                      weights = dataCar$numclaims)
```

```{r}
severity_model
```

## Gamma deviance formula

Deviance formulas can be found here;  http://docs.h2o.ai/h2o/latest-stable/h2o-docs/data-science/glm.html. 

```{r}
gamma_deviance <- function(y, p, w) {
  
  d <- 2 * w * (-log(y / p) + ((y - p) / p))

  dev <- sum(d)
  
  return(dev)
  
}
```

## Compare to results from glm()

Note the model excludes records with no claims so these are excluded when calculating the deviance.

```{r}
clms <- dataCar$numclaims > 0
```

```{r}
gamma_deviance(y = dataCar$average_claimcst0[clms], 
               p = severity_model$fitted.values,
               w = dataCar$numclaims[clms])
```

```{r}
severity_model$deviance
```

```{r}
gamma_deviance(y = dataCar$average_claimcst0[clms], 
               p = sum(dataCar$average_claimcst0[clms] * dataCar$numclaims[clms]) / sum(dataCar$numclaims[clms]),
               w = dataCar$numclaims[clms])
```

```{r}
severity_model$null.deviance
```

# Package info

```{r record_pkg_info}
sessionInfo()
```
















