---
title: "Glm-Important-Variables"
output: github_document
---

This notebook shows how standardising variables when building regression models results in **standardised coefficients** which are independant of the scales of the predictors. This allows the use of regression coefficients as a general measure of variable importance in the model. 

## Start h2o

```{r setup, include = TRUE}
suppressMessages(library(h2o))
h2o.init()
h2o.no_progress()
```

## Upload data to h2o

First upload the prostate dataset from the h2o package to the h2o instance.

```{r load_data}
prostate.hex = h2o.uploadFile(path = system.file("extdata", "prostate.csv",
                                                 package = "h2o"),
                               destination_frame = "prostate.hex")
```

Add another age column where the units are in days. We will build a model using each age variable.

```{r}
prostate.hex["AGE_DAYS"] = prostate.hex["AGE"] * 365
prostate.hex
```

## Build two penalised regression model

We are specifying the `standardize = TRUE` option to standardise numeric variables to have zero mean and unit variance (note it is set to `TRUE` by default). See [h2o docs](http://docs.h2o.ai/h2o/latest-stable/h2o-docs/data-science/algo-params/standardize.html) for more info.

Build the first model using age in years.

```{r build_glm}
explanatory_cols <- c("AGE", "RACE", "DPROS", "DCAPS", "PSA", "VOL", "GLEASON")
h2o_glm <- h2o.glm(x = explanatory_cols,
                   y = "CAPSULE",
                   training_frame = prostate.hex,
                   lambda_search = TRUE,
                   alpha = 1,
                   standardize = TRUE)
```

```{r show_model}
h2o_glm
```

Now build the model using age in days.

```{r build_glm2}
explanatory_cols <- c("AGE_DAYS", "RACE", "DPROS", "DCAPS", "PSA", "VOL", "GLEASON")
h2o_glm2 <- h2o.glm(x = explanatory_cols,
                    y = "CAPSULE",
                    training_frame = prostate.hex,
                    lambda_search = TRUE,
                    alpha = 1,
                    standardize = TRUE)
```

```{r show_model2}
h2o_glm2
```

We see that the standardised coefficients are the same (to several dp) for all variables including age where the scale has changed. For the non-standardised coefficients this is not the case.

## Package info

```{r record_pkg_info}
sessionInfo()
```
